The ChEMBL database as Linked Open Data

### 2013-04-20 article.tex
Removed a confusing line which was not adding new information
Applied the (minor) changes by Janna

However, where the database content is well-defined, specific ontologies have been used.
For example, where the database specifies literature from which data was extracted,
the CiTO ontology is used. The choice of the ontologies is driven by the use cases and
compatibility with other Linked Data projects, rather than formal ontological arguments.
We have not observed ontological contradictions based on these choices.

@@

It should also be noted that all drug-like entities in ChEMBL are not ontologically
subclassing a drug class. Instead, we give the entities the role of drug, and only within a
particular context would they be a drug. If not used as a medication, they are not a drug.
Moreover, the triples only attach this drug role to approved drugs, as defined by the
ChEMBL database. 
% FIXME: What does the following exactly mean? It isn't clear whether it actually means:
% Not all drug-like entities in the database have the role of drug
All drug-like entities in the database do not have the role of drug.
The role is triplified using the OBO and ChEBI ontologies in the following manner,
where CHEBI\_23888 is the ontological entry for ``drug role'' in ChEBI:

@@

Much could be argued about the choice of ontologies, and the lack of an
ontology in places where we decided to use custom classes and predicates. These
choices have only partly been decided using ontological arguments. Most of the choices are based
on pragmatic approaches to enable solutions to the range of use cases we describe here. It should be noted,
however, that the ChEMBL database captures a wide variety of experimental studies
and that semantics are not fully specified in the primary source from which the ChEMBL data was
extracted. As it is not possible to infer the original intentions in these cases, 
it is our opinion that it is better to leave some semantics undefined. During the 
development of the applications some incorrect assumptions have been discovered and
fixed, but in the main we have tried to avoid making assumptions if they are not generally known to be valid. 
Therefore, the goal of this work is the conversion of the ChEMBL data into
RDF, and not an ontological modeling of the domain.

@@

Finally, the ChEMBL team (JO, AG, MD) plans to be a primary source of a RDF version of ChEMBL, while others continue
exposing ChEMBL as RDF too (EW, PA). In the future various versions may arise, exposing the data in different ontologies,
supporting current community needs, however, convergence is expected after a period of experimenting with ontologies that
may not prove to be practically useful. The availability of Open Source scripts ensures that the community can contribute to the
process of choosing a practically useful set of ontological terms.

\section*{Authors contributions}
EW initiated the project, created the initial RDF version of the ChEMBL data, and encouraged the use cases.
AW extended CitedIn to support citation info in ChEMBL-RDF and supported setting up the http://linkedchemistry.info/ host.
OS, AJW, and VT developed the nearest neighbor application of Bioclipse and ChemSpider.
JO, AG, and MD critically validated and summarized the ChEMBL data statistics.
BC, DW contributed the Chem2Bio2RDF use cases.
All authors contributed to the continued development of ChEMBL-RDF, as well as to the writing of the paper,
and approved the final version.

@@

\begin{table*}
\caption{Prefixes and their matching namespaces used in this paper.} \label{namespaces}
\begin{center}
\begin{tabular}{ll}
\hline
\multicolumn{2}{l}{\textbf{Common Vocabularies}} \\
bibo    & Bibliography Ontology~\cite{Giasson2011} \\
        & http://purl.org/ontology/bibo/ \\
chebi   & Chemical Entities of Biological Interest~\cite{DeMatos2010} \\
        & http://purl.org/obo/owl/CHEBI\# \\
cheminf & Chemical Information Ontology~\cite{Hastings2011} \\
        & http://semanticscience.org/resource/ \\
cito    & Citation Typing Ontology~\cite{Shotton2010} \\
        & http://purl.org/spar/cito/ \\
obo / pro & OBO \& PRotein Ontology~\cite{Natale2011} \\
          & http://purl.obolibrary.org/obo/ \\
bfo     & Basic Formal Ontology~\cite{Smith2004} \\

>>>>

However, where the database content is well-defined, specific ontologies have been used.
For example, where the database specifies literature from which data were extracted,
the CiTO ontology is used. The choice of the ontologies is driven by the use cases and
compatibility with other Linked Data projects, rather than formal ontological considerations.
We have not observed ontological contradictions based on these choices.

@@

It should also be noted that all drug-like entities in ChEMBL are not ontologically
subclassing a drug class. Instead, we give the entities the role of drug, and only within a
particular context would they be a drug. If not used as a medication, they are not a drug.
Moreover, the triples only attach this drug role to approved drugs, as defined by the
ChEMBL database.
The role is triplified using the OBO and ChEBI ontologies in the following manner,
where CHEBI\_23888 is the ontological entry for ``drug role'' in ChEBI:

@@

Much could be argued about the choice of ontologies, and the lack of an
ontology in places where we decided to use custom classes and predicates. Most of the choices are based
on pragmatic approaches to enable solutions to the range of use cases we describe here. It should be noted,
however, that the ChEMBL database captures a wide variety of experimental studies
and that semantics are not fully specified in the primary source from which the ChEMBL data was
extracted. As it is not possible to infer the original intentions in these cases, 
it is our opinion that it is better to leave some semantics undefined. During the 
development of the applications some incorrect assumptions have been discovered and
fixed, but in the main we have tried to avoid making assumptions if they are not generally known to be valid. 
Therefore, the goal of this work is the conversion of the ChEMBL data into
RDF, and not an ontological modeling of the domain.

@@

Finally, the ChEMBL project will privide RDF versions of ChEMBL, while others continue
exposing ChEMBL as RDF too (EW, PA). In the future various versions may arise, exposing the data in different ontologies,
supporting current community needs, however, convergence is expected after a period of experimenting with ontologies that
may not prove to be practically useful. The availability of Open Source scripts ensures that the community can contribute to the
process of choosing a practically useful set of ontological terms.

\section*{Authors contributions}
EW initiated the project, created the initial RDF version of the ChEMBL data, and encouraged the use cases.
AW extended CitedIn to support citation info in ChEMBL-RDF and supported setting up the http://linkedchemistry.info/ host.
OS, AJW, and VT developed the nearest neighbor application of Bioclipse and ChemSpider.
BC, DW contributed the Chem2Bio2RDF use cases.
All authors contributed to the continued development of ChEMBL-RDF, as well as to the writing of the paper,
and approved the final version.

@@

\begin{table*}
\caption{Prefixes and their matching namespaces used in this paper.} \label{namespaces}
\begin{center}
\begin{tabular}{ll}
\hline
\multicolumn{2}{l}{\textbf{Common Vocabularies}} \\
bibo    & Bibliography Ontology~\cite{Giasson2011} \\
        & http://purl.org/ontology/bibo/ \\
chebi   & Chemical Entities of Biological Interest~\cite{Hastings2013} \\
        & http://purl.org/obo/owl/CHEBI\# \\
cheminf & Chemical Information Ontology~\cite{Hastings2011} \\
        & http://semanticscience.org/resource/ \\
cito    & Citation Typing Ontology~\cite{Shotton2010} \\
        & http://purl.org/spar/cito/ \\
obo / pro & OBO \& PRotein Ontology~\cite{Natale2011} \\
          & http://purl.obolibrary.org/obo/ \\
bfo     & Basic Formal Ontology~\cite{Smith2004} \\
